### 2024-09-27 paper/paper.qmd
valentines day table

---
title: "Seasonal Variation in Marriage License Applications"
subtitle: "An Exploratory Analysis with Open Data Toronto"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at: https://github.com/fatimahsy/Pandemic-Marriages-.git."
date: today
date-format: long
abstract: "**IN PROGRESS**       
What was done:Marriage License data from Toronto between 2011 and 2024 was analysed to determine whether holidays influence the number of applications.  What was found: Months associated with popular holidays like February (Valentine's day) and December( Christmas), show an increase in marriage license applications. Why this matters:Understanding these patterns can help city officials anticipate changes in demand for marriage-related services and allocate resources accordingly, ensuring that civil centers are prepaered for these seasonal shifts. It also gives us an insight into how emotional and culturl influences during holidays can drive personal descisions.  "
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("readr")
# install.packages("janitor")
# install.packages("opendatatoronto")
# install.packages("knitr")
# install.packages("here")
# install.packages("kableExtra")
# install.packages("lubridate")
library(lubridate)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(janitor)
library(opendatatoronto)
library(knitr)
library(here)
library(kableExtra)

data <- read_csv(here::here("data/analysis_data/analysis_data"))

```

# Introduction

Marriage decisions are often driven but just in terms of practicality, but in terms of deeply emotional and sometimes irrational impulses. Symbolic dates like Valentine's Day and Christmas play a key role in this. Even during unprecedented times like the COVID-19 pandemic, love and tradition lead people to marry at significant moments, despite the logistical nightmares these choices presented. This paper examines trends in marriage licenses issued in Toronto from 2011 to 2023, with a focus on seasonal periods like February and December.

Wile there have been studies showing decreases in marriage licences during global crisis, there has been little analysis on how local trends are influenced by emotionally significant dates. This paper fills that gap by examining the effect of these periods on marriage decisions in Toronto, providing information on how tradition and emotion play a part in these decisions.

We use data analysis and various visualizations to check trends from 2011-2023 with a focus on special seasons. The findings show that while marriage licence declined significantly during the pandemic, Christmas and Valentines Day continued to see small spikes, showing that symbolic dates have some sort of power in marriage decisions.

Marriage trends are a very good indicator of changes in society, reflecting broader shifts in demographics, economic circumstances, and changing cultural preferences. Understanding how external factors like seasonal holidays and different disruptions to the overall society influence marriage decisions can offer us valuable insights into human behavior and societal recovery. By identifying consistent seasonal patterns, this analysis can help wedding-retated businesses (venues, caters, DJ's, etc.)anticipate periods of higher demand and better allocate resource. Additionally, civil centers can use this data to optimise staffing in order to increase efficiency.

The Data @sec-data of this paper explores the `opendatatoronto` data and outlines tools and methods used to analyze the data and some preliminary observations observed. The Results @sec-results introduces more observations found from the data analysis. The Discussion @sec-disc ties back the findings to its real world relevance. Lastly, the Conclusion @sec-concl summarizes key findings and outlines future areas of study.

# Data {#sec-data}

The dataset for this report consist of statistics for marriage licence quantities by civic centre's in Toronto from 2011-2023. This data comes from `opendatatoronto`, a public repository that provides access to different civil datasets. These statistics were downloaded, cleaned, parsed, analysed, and visualised using R (@citeR), a statistical programming language, along with package support from `tidyverse` (@citetidyverse), various different libraries such as:

-   `ggplot2` [@citeggplot]
-   `dplyr` [@citedplyr]
-   `readr` [@citereadr]
-   `tibble` [@citetibble]

For further cleaning, the `janitor` (@citejanitor) package was used and the `knitr` was used too.

## Variables of Interest

```{r}
#| label: marriage-variables
#| fig-cap: "Table 1: Variable Overview"
#| echo: false
#| warning: false
#| message: false
library(tidyverse)
library(here)
library(kableExtra)
data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

# Creating an overview of the data
data_overview <- tibble(
  Variable = c("CIVIC_CENTRE", "MARRIAGE_LICENSES", "TIME_PERIOD", "ID" ),
  Description = c(
    "Code representing the civic center where the marriage license was issued",
    "Number of marriage licenses issued in that month",
    "The year and month when the licenses were issued, formatted as YYYY-MM",
    "A unique identifier for each record"
  ),
  Data_Type = c(
    "Character",  
    "Numeric",
    "Date",
    "Numeric"
  )
)

# Display the table
knitr::kable(data_overview, format = "latex", col.names = c('Variable', 'Description', 'Data Type')) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), full_width = FALSE) %>%
  column_spec(2, width = "25em")

```

The dataset has 544 rows and 6 columns. Each row represents the number of marriage licenses issued at a civic center during a given month, along with a corresponding year and a constructed date column. (@marriage-variables) describes the variables included in the dataset. These variables include:

-   `civic_centre`
-   `marriage_licenses`
-   `time_period`
-   `id`

The data is organised by moth and aggregated for each civic center. While it would have been beneficial to have data at the daily level, the monthly data still allows for meaningful analysis of seasonal trends.

## Data cleaning 
The raw data was cleaned to prepare it for analysis by removing columns, splitting the time information, and creating a date column. The cleaning process standardized the data making it easier to interpret. Here were the steps taken:

-  Rename Columns:Column names were standardized using `clean_names()` for consistency.

- Split time_period into year and month: The time_period column, originally formatted as YYYY-MM, was split into separate year and month columns for easier manipulation.

- Create date column: A new date column was constructed using the `lubridate::ymd()` function.

## Dataset Overview
Table 1 Shows a sample of the dataset used for the analysis of this paper. Here we can see `civic_centre` indicates the name of the civic center where the marriage licenses were issued (e.g., ET for Etobicoke, NY for North York, SC for Scarborough, and TO for Toronto).

In addition to the tabular data, Figure 1 shows the overall distribution of marriage licences by month. The boxplot illustrates the seasonal trends with the number of licences peaking in the summer months, which are particularly wedding months.

```{r}
#| label: Toronto-marriages
#| fig-cap: "Table 2: Sample of Toronto Marriage Licenses"
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

# Load the raw data
raw_data <- read_csv(here::here("data/raw_data/raw_data.csv"))

# Use the head function to show the first few rows
sample_data <- head(raw_data)

# Create a nicely formatted table using kableExtra
sample_data %>%
  kbl(col.names = c("ID", "Civic Centre", "Marriage Licenses", "Time Period"), 
      caption = "Sample of Toronto Marriage Licenses") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = F)
```

```{r}
#| label: distribution-by-month
#| fig-cap: "Figure 1: Distribution of Marriage Licenses by Month"
#| echo: false
#| warning: false
#| message: false
library(ggplot2)
library(dplyr)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# Box Plot: Distribution of Marriage Licenses by Month
ggplot(data, aes(x = as.factor(month), y = marriage_licenses)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Distribution of Marriage Licenses by Month",
       x = "Month",
       y = "Number of Licenses") +
  theme_minimal()

```

```{r}
#| label: total-by-civic-centre
#| fig-cap: "Figure 2: Total Marriage Licenses by Civic Centre"
#| echo: false
#| warning: false
#| message: false

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

# Bar Plot: Total Marriage Licenses by Civic Centre
ggplot(data, aes(x = civic_centre, y = marriage_licenses, fill = civic_centre)) +
  stat_summary(fun = sum, geom = "bar") +
  labs(title = "Total Marriage Licenses by Civic Centre",
       x = "Civic Centre",
       y = "Total Number of Licenses") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") #adding a little pazazz!
```
Figure 2 shows a breakdown of the total marriage licenses issued by each center from 2011-2023. Toronto issued the highest number of licences with over 90,000 licenses across the period, followed by North York, Scarborough and Etobicoke. The major disparity in marriage license counts across centers can mainly be attributed to differences in population density and geographical location. Toronto as the central and most densely populated area, attracts the largest number of couples applying for marriage licenses.

```{r}
#| label: fig-marriage
#| fig-cap: Marriage Licinces By Date
#| echo: false

data |> 
  ggplot(aes(x = date, y = marriage_licenses)) +
  geom_point()

```
The scatter plot (@fig-marriage) displays the number of marriage licenses issued in Toronto over time, spanning from before 2011 to approximately 2025.Each point represents the number of liciences issued in a specific month. the scatter plot illustrates both the overall distribution of licenses over time and any significant deviations or disruptions to typical patterns.

# Results {#sec-results}

## Licenses by Year and Month
```{r}
#| label: licenses-heatmap
#| fig-cap: "Figure 4: Heatmap of Marriage Licenses by Year and Month"
#| echo: false
#| warning: false
#| message: false

# Heatmap: Marriage Licenses by Year and Month
data %>%
  mutate(month_year = format(date, "%Y-%m")) %>%
  ggplot(aes(x = year, y = as.factor(month), fill = marriage_licenses)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "steelblue") +
  labs(title = "Heatmap of Marriage Licenses by Year and Month", 
       x = "Year", 
       y = "Month") +
  theme_minimal()

```
Figure 4 shows a heatmap of the distribution of marriage licenses issued by month and year from 2022 to 2024. The x-axis represents years and the y-axis represents the months. The colour gradient illustrates the number of marriage licenses issued, with the darker shades representing more licenses issued. Higher volumes of marriage licenses are typically observed during the summer months(June-August) across all years, as shown by the darker shades on the heatmap. Winter months tend to have fewer marriage licenses issued. In 2020, there was a clear drop in the number of licences, particularly during the months that typically see the most activity.

## Valentines Day Licenses

Table 3 shows the total number of marriage licenses issued in February for each year between 2011-2024.Contrary to my earlier belief, February is typically a low-activity month for marriage license, with most years seeing less than 900 licenses issued. However, in 2020, 908 licenses were issued in February, the ost in the whole period. They year 2021 shows a significant drop in licenses, with only 535 licenses issued, reflecting the ongoing impact of the COVID-19 pandemic. By 2024, the number of february licenses reached 993, the highest for any February in the dataset, indicating a recovery and potential resurgence in wedding activity during this month.

```{r}
#| label: february-table
#| fig-cap: "Total Marriage Licenses Issued in February for Each Year"
#| echo: false
#| warning: false
#| message: false

# Summarize February Data: Filter for February and calculate the total number of licenses
february_data <- data %>%
  filter(month == "02") %>%  # Filter for February
  group_by(year) %>%  # Group by year
  summarise(february_licenses = sum(marriage_licenses))  # Summarize total licenses in February

# Create a table of February marriage licenses for each year
knitr::kable(
  february_data,
  caption = "Table: Total Marriage Licenses Issued in February for Each Year",
  col.names = c("Year", "February Licenses"),
  format = "simple"
)

```

## Pre,During & Post Pandemic Licenses
```{r}
#| label: avg-licenses-bar
#| fig-cap: "Figure: Average Monthly Marriage Licenses Pre-, During-, and Post-Pandemic"
#| echo: false
#| warning: false
#| message: false

# Group data by period (pre-pandemic, pandemic, post-pandemic) and calculate average monthly licenses
total_licenses_period <- data %>%
  mutate(period = case_when(
    date < as.Date("2020-03-01") ~ "Pre-Pandemic",
    date >= as.Date("2020-03-01") & date <= as.Date("2021-12-31") ~ "Pandemic",
    date > as.Date("2021-12-31") ~ "Post-Pandemic"
  )) %>%
  group_by(period) %>%
  summarise(avg_monthly_licenses = mean(marriage_licenses))

# Bar Plot: Average Monthly Marriage Licenses by Period
ggplot(total_licenses_period, aes(x = period, y = avg_monthly_licenses, fill = period)) +
  geom_bar(stat = "identity", width = 0.6) +
  labs(title = "Average Monthly Marriage Licenses Pre-, During-, and Post-Pandemic", 
       x = "Period", 
       y = "Average Monthly Licenses") +
  theme_minimal() +
  scale_fill_manual(values = c("Pre-Pandemic" = "blue", "Pandemic" = "red", "Post-Pandemic" = "green"))

```
Figure 5 shows the average number of licenses issued per month during three distinct periods: pre-pandemic, pandemic & post-pandemic.Before the pandemic(2011-2019),there was an average of approximately 300 licenses issued per mont. During the pandemic, the red bar on the graph shows a significant drop in the average monthly licenses with approximately 450 licenses issued monthly. This reflects the overall decline in marriage due pandemic-related restrictions and disruptions. The green bar represents the post-pandemic period, showing an increase in marriage licenses issued with an average of 700 licenses per month. This suggest a rebound in normal activities. This hart clearly shows the disruption caused by the pandemic as well as later recovery in the following years, with post-pandemic averages exceeding pre-pandemic averages.

## December Licenses
```{r}
#| label: december-table
#| fig-cap: "Table 3: Total Marriage Licenses Issued in December for Each Year"
#| echo: false
#| warning: false
#| message: false

# Summarize December Data: Filter for December and calculate the total number of licenses
december_data <- data %>%
  filter(month == "12") %>%  # Filter for December
  group_by(year) %>%  # Group by year
  summarise(december_licenses = sum(marriage_licenses))  # Summarize total licenses in December

# Create a table of December marriage licenses for each year
knitr::kable(
  december_data,
  caption = "Table: Total Marriage Licenses Issued in December for Each Year",
  col.names = c("Year", "December Licenses"),
  format = "simple"
)

```

```{r}
#| label: december-plot
#| fig-cap: "Figure 2: Total Marriage Licenses Issued in December (2011-2023)"
#| echo: false
#| warning: false
#| message: false

# Plot Total December Licenses Over Time
ggplot(december_data, aes(x = year, y = december_licenses)) +
  geom_line(color = "green", size = 1) +
  geom_point(color = "red", size = 3) +  # Optional: Add points for each year
  labs(title = "Total Marriage Licenses Issued in December (2011-2023)",
       x = "Year",
       y = "Number of Licenses") +
  theme_minimal()

```

## Patterns

# Discussion {#sec-disc}

# Conclusion {#sec-concl}

\`\`\`

\newpage

\appendix

# Appendix {.unnumbered}
```{r}
#| label: monthly-totals-table
#| fig-cap: "Total Monthly Marriage Licenses by Year"
#| echo: false
#| warning: false
#| message: false

# Monthly Summary for All Months: Group data by year and month to calculate total licenses
monthly_totals <- data %>%
  group_by(year, month) %>%
  summarise(total_licenses = sum(marriage_licenses))

# Create a table of monthly totals by year
knitr::kable(
  tidyr::spread(monthly_totals, year, total_licenses),  # Reshape data to display years side by side
  caption = "Table: Total Monthly Marriage Licenses by Year",
  col.names = c("Month", unique(monthly_totals$year)),  # Dynamically add years as column names
  format = "simple"
)

```
# References